Smoller RUV-2 
========================================================

```{r setup, echo=FALSE}
opts_chunk$set(tidy=TRUE, echo=TRUE, highlight=TRUE, figalign='center', fig.height=9, fig.width=9, out.width='800px', message=FALSE, error=TRUE, warning=FALSE, cache=FALSE)

# Setup report details
clientname="Erin Dunn"
clientemail="erindunn@pngu.mgh.harvard.edu"
lablocation="MGH"
analystname="Meeta Mistry"
analystemail="mmistry@hsph.harvard.edu"
```

Array analysis for `r clientname` (`r clientemail`) at `r lablocation`. Contact `r analystname` (`r analystemail`) for additional details. Request from client was:

Test for differential gene expression between brain samples taken at different ages as part of the [BrainCloud](http://braincloud.jhmi.edu/BrainCloudHelp.htm) project. 

> RNA from 269 human prefrontal cortex samples ranging from fetal development (negative ages) through aging (80 years) were analyzed on custom 2-color microarrays from the National Human Genome Research Institute (NHGRI) microarray core facility using a reference RNA comprised of a pool of all samples.

# Workflow
Try starting from the pre-'cleaned' data and remove the effects of confounders using [RUV-2](http://www.stat.berkeley.edu/~johann/ruv/).

## Setup

### Bioconductor and R libraries used

```{r libraries, echo=TRUE}
library(ggplot2)
library(gtable)
library(scales)
library(GGally)
library(RColorBrewer)
library(GEOquery)
library(affy)
library(arrayQualityMetrics)
library(hgu133plus2.db)
library(reshape)
library(xtable)
library(ruv)
library(limma)
library(Biobase)

# Grab John Hutchinson's convenience functions
source("http://dl.dropboxusercontent.com/u/4253254/Resources/functions.r")
```

```{r variables, echo=TRUE}
# Setup Data and Results directory variables
baseDir <- '.'
dataDir <- file.path(baseDir, "data")
metaDir <- file.path(dataDir, "meta")
resultsDir <- file.path(baseDir, "results")
#covarsfilename <- 'covdesc.txt'
```

### Load the data and metadata

```{r dataimport GEO}

# Load GEO data
gpl <- getGEO('GSE30272', destdir=file.path(dataDir, 'geo'))
eset<-as(gpl$GSE30272_series_matrix.txt.gz, "ExpressionSet")

# Get expression data
expression<-exprs(eset)
colnames(expression)<-pData(eset)$title
dim(epxression)

```


```{r metadataImport}
meta <- read.table(file.path(metaDir, 'brain_pheno2-txt.tsv'),
                   header=TRUE, sep='\t', na.strings='NULL', row.names=1)

# Remove outlier samples
remove<-c(which(meta$Race == "HISP"), which (meta$Race == "AS"), which(meta$Gender == 5))
newMeta<-meta[-remove,]
newMeta<-newMeta[order(newMeta$Age),]

# Align metadata order with expression data
expression<-expression[,which(colnames(expression) %in% newMeta$geo_accession)]
rownames(newMeta)<-newMeta$geo_accession
newMeta<-newMeta[,-1]
newOrder <- match(colnames(expression), rownames(newMeta))
expression.meta <- newMeta[as.vector(newOrder), ]
all(colnames(expression) == rownames(expression.meta))

```

```{r setup for RUV}

pData(eset)<-newMeta
mod <- model.matrix(~Age -1, pData(eset))
X<-as.matrix(mod[,1])
Y<-t(expression)

# Get gene annotations
gpl <- getGEO('GPL4611', destdir=file.path(dataDir, 'geo'))
annot <- Table(gpl)[, c('ID', 'Gene_Symbol', 'Entrez_Gene_ID')]

rownames(annot)<-annot[,1]
annot<-annot[,-1]
```

